func-descr.template

### 2025-12-21 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd

Below is the [lines of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1578) that determines the optimal range of `span` for the further model tuning when `degree = 0`:
```{r, eval=FALSE}
  spans <- seq(0.0005, 1, 0.001)
  lss.UMGYDE.preset.degree0.result <- 
    tune.model_param(spans, train_UMGY_SmoothedDay_effect.RMSE.cv.degree0)
```
```{r}
str(lss.UMGYDE.preset.degree0.result)
```

@@

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYDE Model_. For the current model, this formula takes the form:

@@

And the formula \@ref(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section is available in the [UMGYDE Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2322) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2322) script on _GitHub_.
:::

Let's perform the preconfiguration to determine the appropriate range of $\lambda$ for subsequent fine-tuning of our current model:

We are going to use the [tune.model_param](#func.tune.model_param) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of [Appendix A](#appndx_a), passing the [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv) function as the value of the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) parameter.

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
  lambdas <- seq(0, 256, 16)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

Now, let's visualize the results of the $\lambda$ range preconfiguration:
``` {r}
cv.UMGYDE.preset.result$tuned.result |>
  data.plot(title = TeX(r'[UMGYDE Model Regularization: $\lambda$ Range Pre-configuration]'),
              xname = "parameter.value", 
              yname = "RMSE", 
              xlabel = TeX(r'[$\lambda$]'), 
              ylabel = "RMSE")
```

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of [Appendix A](#appndx_a), which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section can be found in the [UMGYDE Model Regularization: Fine-tuning](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2378) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2378) script on _GitHub_.
:::

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UMGYDE.preset.result$tuned.result)

UMGYDE.loop_starter <- c(endpoints["start"], 
                        endpoints["end"], 
                        8)

cache.base_name <- "UMGYDE.rglr.fine-tuning"

UMGYDE.rglr.fine_tune.results <- 
  model.tune.param_range(UMGYDE.loop_starter,
                         UMGYDE.rglr.fine_tune.cache.path,
                         cache.base_name,
                         regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

::: {.noteblock data-latex=""}
Under the hood, this code block internally calls the helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of [Appendix A](#appndx_a) to figure out the boundaries of the range of values from the nearest neighborhood of the value corresponding to the best `RMSE` in the data passed in as a parameter.
:::

@@

@@

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of [Appendix A](#appndx_a), which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

>>>>

The following [piece of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1578) determines the optimal range of `span` for the further model tuning when `degree = 0`:
```{r, eval=FALSE}
  spans <- seq(0.0005, 1, 0.001)
  lss.UMGYDE.preset.degree0.result <- 
    tune.model_param(spans, train_UMGY_SmoothedDay_effect.RMSE.cv.degree0)
```
```{r}
str(lss.UMGYDE.preset.degree0.result)
```

@@

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be found in the [UMGYDE Model Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1239) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L1239) script.
:::

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:YE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

@@

And the formula \@ref(eq:YE-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section is available in the [UMGYDE Model Regularization: Pre-configuration](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2053) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2053) script on _GitHub_.
:::

As for previous models, we will use the function [tune.model_param](#func.tune.model_param) passing in the [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value) argument the model-specific helper function (this time designed for the *UMGYDE Model*): [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv).

::: {.noteblock data-latex=""}
The functions [tune.model_param](#func.tune.model_param) and [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv) are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

The following [piece of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2070) performs this operation:
```{r, eval=FALSE}
  lambdas <- seq(0, 256, 16)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

Now, let's visualize the results of the $\lambda$ range pre-configuration:
``` {r}
cv.UMGYDE.preset.result$tuned.result |>
  data.plot(title = TeX(r'[UMGYDE Model Regularization: $\lambda$ Range Pre-configuration]'),
              xname = "parameter.value", 
              yname = "RMSE", 
              xlabel = TeX(r'[$\lambda$]'), 
              ylabel = "RMSE")
```

@@

::: {.noteblock data-latex=""}
The complete version of the source code shown in this section can be found in the [UMGYDE Model Regularization: Fine-tuning](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2109) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2109) script on _GitHub_.
:::

@@

The following [code snippet](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2110) prepares the interval of values for the $\lambda$ parameter, over which the operation has to be done:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UMGYDE.preset.result$tuned.result)

UMGYDE.loop_starter <- c(endpoints["start"], 
                        endpoints["end"], 
                        8)
```
```{r}
UMGYDE.loop_starter
```

::: {.noteblock data-latex=""}
The helper function [get_fine_tune.param.endpoints](#func.get_fine_tune.param.endpoints) used in code snippet above is described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) of [Appendix A: Support Functions].
:::

And the next [piece of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L2120) below accomplishes the task of *fine-tuning* the model:
```{r, eval=FALSE}
cache.base_name <- "UMGYDE.rglr.fine-tuning"

UMGYDE.rglr.fine_tune.results <- 
  model.tune.param_range(UMGYDE.loop_starter,
                         UMGYDE.rglr.fine_tune.cache.path,
                         cache.base_name,
                         regularize.test_lambda.UMGYD_effect.cv)

UMGYDE.rglr.fine_tune.RMSE.best <- UMGYDE.rglr.fine_tune.results$best_result["best_RMSE"]
```

::: {.noteblock data-latex=""}
The custom functions [model.tune.param_range](#func.model.tune.param_range) and [regularize.test_lambda.UMGYD_effect.cv](#func.regularize.test_lambda.UMGYD_effect.cv) used in the code snippet above are described in the sections *Common Helper Functions /* [Model Tuning Utils](#appndx_a.CHF.model_tuning_utils) and  *Models Training: Support Functions /* [UMGYE Model: Regularization], respectively, of [Appendix A: Support Functions].
:::

@@

Below are the results of fine-tuning the *UMGYDE Model*:

```{r, echo=FALSE}
writeLines("*** Path to the cache directory for intermediate fine-tuning results ***")
```
```{r}
UMGYDE.rglr.fine_tune.cache.path
```
```{r, echo=FALSE}
writeLines("*** Fine-tuning results object data structure ***")
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
```
```{r, echo=FALSE}
writeLines("*** Fine-tuning: best results ***")
```
```{r}
UMGYDE.rglr.fine_tune.results$best_result
UMGYDE.rglr.fine_tune.RMSE.best
```
\newpage

@@

::: {.noteblock data-latex=""}
Note that in the code snippet above, we use the custom data visualization function [data.plot](#func.data.plot) (described in the *Data Helper Functions /* [Data Visualization Functions] section of [Appendix A](#appndx_a)) with the argument [normalize](#func.data.plot.args.normalize) set to `TRUE`, which means that deviations from the minimum $y$ value are used to plot, rather than the $y$ values themselves.
:::

### 2025-12-21 reports/capstone-movielens.site/work/draft-snippert.Rmd

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Common Helper Functions: Regularization](#appndx_a.CHF.regularization) section of [Appendix A: Support Functions]).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

@@

### Regularization Functions {#appndx_a.CHF.regularization}
\

@@

### Regularization Functions2 {#appndx_a.CHF.regularization2}
\

>>>>

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the Section [Common Helper Functions: Regularization](#appndx_a.CHF.model_tuning_utils) section of [Appendix A: Support Functions]).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

@@

### Regularization Functions {#appndx_a.CHF.model_tuning_utils}
\

@@

### Regularization Functions2 {#appndx_a.CHF.model_tuning_utils2}
\
